Figures for Paper

### 2021-08-05 paper_figures.Rmd
Update paper_figures.Rmd
added plot for 2F and S2C

```
# NMD experiment on SH-SY5Y cells

@@

fbp_sy <- read_csv(file.path(here::here(),"data","results_17022021_sy5y.csv"))
fbp_sy$condition <- as.factor(fbp_sy$condition)
fbp_sy$condition <- factor(fbp_sy$condition,levels(fbp_sy$condition)[c(2,1)])

stattest_sy <- fbp_sy %>%
    group_by(gene) %>%
    t_test(value ~ 1, mu = 1) %>%
    add_significance() %>%
    add_xy_position(x="gene")
stattest_sy

nmd_plot_sy <- ggbarplot(fbp_sy,
                         x = 'gene',
                         y = 'value',
                         add = c("mean_se","jitter"),
                         fill = 'condition',
                         color = 'condition',
                         position = position_dodge(0.8),
                         dot.size = 10) +
    scale_y_continuous(labels = scales::percent) +
    ggpubr::theme_pubr() +
    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) +
    scale_color_manual(values = c("#1C2617","#262114")) +
    ylab("Percent of transcript after CHX treatment") +
    xlab("") +
    stat_pvalue_manual(stattest_sy, label="p.signif", y.position = 9)
nmd_plot_sy

@@

@@

fbp_dz <- read_csv(file.path(here::here(),"data","results_17022021_dz.csv"))
fbp_dz$condition <- as.factor(fbp_dz$condition)
fbp_dz$condition <- factor(fbp_dz$condition,levels(fbp_dz$condition)[c(2,1)])

stattest_dz <- fbp_dz %>%
    group_by(gene) %>%
    t_test(value ~ 1, mu = 1) %>%
    add_significance() %>%
    add_xy_position(x="gene")
stattest_dz

nmd_plot_dz <- ggbarplot(fbp_dz,
                         x = 'gene',
                         y = 'value',
                         add = c("mean_se","jitter"),
                         fill = 'condition',
                         color = 'condition',
                         position = position_dodge(0.8),
                         dot.size = 10) +
    scale_y_continuous(labels = scales::percent) +
    ggpubr::theme_pubr() +
    scale_fill_manual(values = c("#40B0A6","#E1BE6A")) +
    scale_color_manual(values = c("#1C2617","#262114")) +
    ylab("Percent of transcript after CHX treatment") +
    xlab("") +
    stat_pvalue_manual(stattest_dz, label="p.signif", y.position = 26)
nmd_plot_dz

>>>>

```
# NMD experiment on SH-SY5Y cells - CHX

@@

for_bar_plot_input <- fread(file.path(here::here(),"data","SY5Y_Tidy_4.csv"))
for_bar_plot <- data.frame(for_bar_plot_input)
for_bar_plot$Treatment <- factor(for_bar_plot$Treatment, levels = c("DMSO", "CHX"))

shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "HNRNPL")$Value)
shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "STMN2")$Value)
shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "UNC13A")$Value)
shapiro.test(filter(for_bar_plot, Treatment == "CHX" & Gene == "UNC13B")$Value)

stattest <- for_bar_plot %>%
  group_by(Gene) %>%
  pairwise_t_test(Value ~ Treatment, ref.group = "DMSO", alternative = "greater") %>%
  add_significance() %>% 
  add_xy_position(x="Treatment")
stattest

nmd_plot_chx = ggbarplot(for_bar_plot,
                     x = 'Gene',
                     add = c("mean_se","jitter"),
                     y = 'Value',
                     color = 'Treatment',
                     position = position_dodge(0.8),
                     dot.size = 10) +
  scale_y_continuous() + 
  ggpubr::theme_pubr() + stat_pvalue_manual(stattest, label="p.signif" , x = "Gene")
plot(nmd_plot_chx)

@@

# NMD experiment on SH-SY5Y cells - sTDP43+sUPF1

@@

for_bar_plot_sy_input <- fread(file.path(here::here(),"data","results_upf1.csv"))

for_bar_plot_sy_df <- data.frame(for_bar_plot_sy_input)
for_bar_plot_sy_df$Condition <- as.factor(for_bar_plot_sy_df$Condition)

#assess TDP-43 and UPF1 KD
for_bar_plot_sz <- for_bar_plot_sy_df %>%
  filter(Gene == "TDP43" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1" | Condition == "Non-treated"))

shapiro.test(filter(for_bar_plot_sz, Condition == "siTDP43+siCTRL" & Gene == "TDP43")$Value)
shapiro.test(filter(for_bar_plot_sz, Condition == "siTDP43+siUPF1" & Gene == "TDP43")$Value)

stattest_sz <- for_bar_plot_sz %>%
  pairwise_t_test(Value ~ Condition, ref.group = "Non-treated", alternative = "less") %>%
  add_significance() %>% 
  add_xy_position(x="Condition")
stattest_sz

for_bar_plot_sx <- for_bar_plot_sy_df %>%
  filter(Gene == "UPF1" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1"))

shapiro.test(filter(for_bar_plot_sx, Condition == "siTDP43+siUPF1" & Gene == "UPF1")$Value)

stattest_sx <- for_bar_plot_sx %>%
  pairwise_t_test(Value ~ Condition, ref.group = "siTDP43+siCTRL", alternative = "less") %>%
  add_significance() %>% 
  add_xy_position(x="Condition")
stattest_sx

for_bar_plot_szx <- for_bar_plot_sy_df %>%
  filter(Gene == "TDP43" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1" | Condition == "Non-treated")
         | Gene == "UPF1" & (Condition == "siTDP43+siCTRL" | Condition == "siTDP43+siUPF1"))

#fix position of stats for plot
stattest_sx$xmin[1] <- 1.8
stattest_sx$xmax[1] <- 2.2
stattest_sx$y.position[1] <- 1.3
stattest_sz$xmin[1] <- 0.75
stattest_sz$xmin[2] <- 0.75
stattest_sz$xmax[1] <- 1
stattest_sz$xmax[2] <- 1.25
stattest_sz$y.position[1] <- 1.2
stattest_sz$y.position[2] <- 1.3

nmd_plotz = ggbarplot(for_bar_plot_szx,
                     x = 'Gene',
                     add = c("mean_se","jitter"),
                     y = 'Value',
                     color = 'Condition',
                     position = position_dodge(0.8),
                     dot.size = 10) +
  scale_y_continuous() + ylab("Percent of transcript after UPF1 knockdown") + xlab("Gene") + 
  ggpubr::theme_pubr() + stat_pvalue_manual(stattest_sz) + stat_pvalue_manual(stattest_sx)
plot(nmd_plotz)

#assess rescue of NMD-sensitive transcripts after UPF1 inhibition
for_bar_plot_sy <- for_bar_plot_sy_df %>%
  filter(Gene == "hnRNPL" | Gene == "STMN2" | Gene == "UNC13A" | Gene == "UNC13B")

shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "hnRNPL")$Value)
shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "STMN2")$Value)
shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "UNC13A")$Value)
shapiro.test(filter(for_bar_plot_sy, Condition == "siTDP43+siUPF1" & Gene == "UNC13B")$Value)

stattest_sy <- for_bar_plot_sy %>%
  group_by(Gene) %>%
  pairwise_t_test(Value ~ Condition, ref.group = "siTDP43+siCTRL", alternative = "greater") %>%
  add_significance() %>% 
  add_xy_position(x="Condition")
stattest_sy

nmd_plot = ggbarplot(for_bar_plot_sy, 
                     x = 'Gene', 
                     add = c("mean_se","jitter"), 
                     y = 'Value',
                     color = 'Condition', 
                     position = position_dodge(0.8), 
                     dot.size = 10) +
  scale_y_continuous() + 
  ggpubr::theme_pubr() + ylab("Percent of transcript after UPF1 knockdown") + xlab("Gene") +
  stat_pvalue_manual(stattest_sy, label="p.signif", x = "Gene")
plot(nmd_plot)
